### 2010-10-21 papers/BMD2010/PhysicalParametersPaper/paper.tex
paper final draft for conference

\begin{center}
\begin{tabular}{c}
$^*$ Department of Mechanical and Aerospace Engineering\\
University of California, Davis\\
One Shields Avenue, Davis, CA, 95616, USA\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu\\
\end{tabular}
\begin{tabular}{c}
$^\dag$ Laboratory for Engineering Mechanics\\
Delft University of Technology\\
Mekelweg 2, 2628CD Delft, The Netherlands\\
e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com\\
\end{tabular} \\ \vspace{2.5ex}
\end{center}
\section*{ABSTRACT}
Accurate measurements of a bicycle's physical parameters are required for
realistic dynamic simulations and analysis. The most basic models require the
geometry, mass, mass location and mass distributions for the rigid bodies. More
complex models require estimates of tire characteristics, human
characteristics, friction, stiffness, damping, etc. In this
paper we present the measurement of the minimal bicycle parameters required for
the benchmark Whipple bicycle model presented in~\cite{Meijaard2007}. This
model is composed of four rigid bodies, has ideal rolling and frictionless joints,
and is laterally symmetric. A set of 25 parameters is used to describe the
geometry, mass, mass location and mass distribution of each of the rigid
bodies. The experimental methods used to estimate the parameters described
herein are based primarily on the work done in~\cite{Kooijman2006} but have
been refined for improved accuracy and methodology. Koojiman's work was
preceded by \cite{Roland1971} who measured a bicycle in a
similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
similar techniques with scooters.

@@

These eight different parameter sets can be used with, but not
limited to, the benchmark bicycle model. The accuracy of all the measurements
are presented up through the eigenvalue prediction of the linear model. The
accuracies are based on error propagation theory with correlations taken into
account.

@@

\section{INTRODUCTION}
This work is intended to document the indirect measurement of eight real
bicycles' physical parameters. The physical parameters measured are those that
are needed for the benchmark Whipple bicycle model presented in
\cite{Meijaard2007}. The work is based on the techiques used to measure the
instrumented bicycles in \cite{Kooijman2006}, \cite{Kooijman2008} and
\cite{Moore2009a}. We improve upon these methods by both increasing and reporting
the accuracies of the measurements and by measuring the complete moments of
inertia of the frame and fork needed for anaylsis of the nonlinear model.
Furthermore, very little data exists on the physical parameters of different
types of bicycles and this work aims to provide a small sample of bicycles.

D\"{o}hring measured the physical parameters of a scooter~\cite{Dohring1953}.
Singh and Goel measured the physical parameters of a scooter~\cite{Singh1971}.
Roland and Massing measured the physical parameters of a bicycle in much the
same was as presented, including calculations of uncertainty from the indirect
measurement techniques~\cite{Roland1971}. Patterson used a swing to measure the
inertia of a bicycle and rider~\cite{Patterson2004}. This work is based off of the work
done by Kooijman~\cite{Kooijman2006}. I worked with Jodi Kooijman for a year and
was able to use much of the same apparatus and refine the measurement
technique.

@@

Where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
matrices are combinations of 25 of the bicycle's physical parameters that

@@

necessarily a minimun set for the Whipple model, as shown in \cite{Sharp2008},
but are useful as they represent more easily measurable quantities.

@@

Figs.~\ref{fig:bicycles}. The three Batavus bicycles were donated by the

@@

variety, a mountatin and road bike. The yellow bike is used to demonstrate

@@

    \item[Batavus Browser (B, B*)]{The Batavus Browser Fig.~\ref{fig:browser} is an averaged priced Dutch city

@@

        bars for an upright posture and includes various accerories for utility
        puposes. We measured the physical proprieties of the stock Browser model and

@@

        Dutch city bike that is setup for touring. It has an aluminum frame, an
        eight speed internal hub, upright handlebars, accesories for utility

@@

        lightweight front suspended mountain bike. It has a ?? frame, large low
        pressure mountain bike tires, and is setup for racing with few extra accerories.}

@@

        made of aluminum and it has a seven speed internal hub. It has mountain

@@

        accesories for utility such as a rear rack, fenders, light and

@@

    \item[Yellow Bicycle (Y, Y*)]{The yellow bicycle Fig.~\ref{fig:yellow} used in the lab to

@@

        aluminum road frame of unknown make with the majority of components

@@

        are the only parts on the bike. The bicycle was measured with both the
        fork in normal position and the fork reversed. The fork was reversed to

@@

The 25 parameters can be estimated by many techniques. Where possible we

@@

We took great care too improve upon the accuracy and report the accuracy of the
measurements of the bicycle parameters. Following the vain of

@@

the measurments, $u,v,\ldots$, then $x$ is a random variable defined as

@@

Using the definitions for variance and covariance

@@

hereafter have uncorrelate variables but a few do not and the covariance has to

@@

The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were

@@

tape measure (resolution: 2mm) was pulled taught and taped on a flat level smooth floor. The tire

@@

$\pm0.2^{\circ}$ accuracy. The bicycle frame was fixed perdenpicular to the
ground, the steering angle was set to the nominal, the tire pressures were are

@@

fork blades where aligned such that the axle axis was parallel to the table

@@

We measured the wheelbase when the bicycle was in the same nominal
configuration as described in Sec.~\ref{sec:headtube}. We used a tape measure

@@

    \caption{\subref{fig:level} The digital level was mounted to a straight edge that was aligned

@@

orientations~\ref{fig:triangle}. The slope can be shown to be

@@

Only two lines are required to calculate the center of mass of the laterally

@@

Alternatively the three lines can be treated as an over determined linear

@@

solution is not at the centroid of the triangle made by the intersecting lines.
I am not sure which solution is theoretically the correct one for the location
of the center of mass.

@@

@@

compound and torsional pendulums and measuring their period of oscillation when
perturb at small angles. The rate of oscillation was measured

@@

\textsc{Matlab}. The measurement duration were either 15 or 30 secs and each moment of

@@

determine only the period of oscillation which is need for the moment of inertia

@@

The following function Eqn~\ref{eqn:decayOs} was fit to the data using a nonlinear least squares fit
routine for each experiment.

@@

The physical phenomena observed for corresponding to data sets such as these were that
the bicycle frame or fork was perturbed torsionally. After set into motion the
torsional motion died out and a longitudinal swinging motion increased. The
motions alternated back and forth with neither ever reaching zero. The
frequencies of these motions were very close to one another and it is not
apparent how dissect the two. We explored fitting to a function such as
    f(t) = A\sin{(\omega_1 t)} + B\sin{(\omega_2 t + \phi)} + C
    \label{eqn:sumSines}
But the fit predicts that $\omega_1$ and $\omega_2$ have very similar
frequencies. There was no easy way to choose which of the two $\omega$'s was
the one associated with the torsional oscillation. Some work was done to model
the torsional pendulum as a laterally flexible beam to determine this, but we
thought the improvement in the accuracy of the period calculation would not
improve enough for the effort required. Future experiments should simply
prevent the swinging motion of the pendulum without damping the torsional
motion.

The period for a damped oscillation is
\begin{equation}
    T = \frac{2\pi}{\sqrt{1-\zeta^2}\omega_n}
    \label{eqn:periodDamped}
\end{equation}
and for light damping
\begin{equation}
    T \approx \frac{2\pi}{\omega_n}
    \label{eqn:period}
\end{equation}
The uncertainty in the period, $T$, can be determined from the fit. Firstly,
the variance of the fit is
\begin{equation}
    \sigma_y^2 =
    \frac{1}{N-5}\sum_{i=1}^N(y_{mi}-\bar{y}_m)^2-(y_{pi}-\bar{y}_m)^2
    \label{eqn:fitVariance}
\end{equation}
The covariance matrix of the fit function can be formed
\begin{equation}
    \mathbf{U} = \sigma_y^2\mathbf{H}^{-1}
    \label{eqn:covariance}
\end{equation}
where $\mathbf{H}$ is the Hessian~\cite{Hubbard1989b}. $\mathbf{U}$ is a $5\times5$ matrix with the
variances of each of the five fit parameters along the diagonal.
The variance of $T$ can be computed using the variance of $\zeta$ and $\omega$. It
is important to note that the uncertainties in the period are very low
($<10e-4$), even for the fits with low $r^2$ values.
\subsection{TORSIONAL PENDULUM}
A torsional pendulum was used to measure all of moments of inertia about axes
in the lateral symmetric plane of each of the wheels, fork and frame. The
pendulum is made up of a rigid mount, an upper clamp, a torsion rod, and
various lower clamps.
\begin{figure}[tbp]
    \begin{center}
        \includegraphics[width=2in]{../../../images/fixture.jpg}
    \end{center}
    \caption{The rigid pendulum fixture mounted to a concrete column.}
    \label{fig:fixture}
\end{figure}
A 5 mm diameter, 1 m long mild steel rod was used as the torsion
spring. A lightweight, low relative moment of inertia clamp was constructed
that could clamp the rim and the tire. The moments of inertia of the clamps
were neglected. The wheel was hung freely such that the
center of mass aligned with the torsional pendulum axis and then it was
secured. The wheel was then perturbed and it oscillated about the pendulum
axis. The rate gyro was mounted on the clamp orientated along the pendulum
axis.

@@

inertia~\ref{fig:rod}. A torsional pendulum almost identical to the one used in
\cite{Kooijman2006} was used to measure the averaged period $\overline{T}_i$ of

@@

are assumed symmetric about three orthogonal planes so there are no products of
inertia. The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring

@@

torsional pendulum setup and Eq.~\ref{eq:torPend}.The wheels are assumed to be symmetric laterally and about any radial axis. Thus only two
moment's of inertia are required for the benchmark parameters set. The moment

@@

gyro was attached to the spokes near the hub and orientated mostly along the

@@

axis whixh added a very low frequency component of rate along the vertical
radial axis, but should not affect the period estimation about the compound
pendulum axis. A fixture that prevented the precession would be preferable for

@@

inertia about the axle is calculated by:

@@

    need the sigma for IRyy

@@

the torsional pendulum axis and then it was secured in position. The wheel was then
perturbed and it oscillated about the pendulum axis. The radial moment of inertia can can calculated as such:

@@

    need the sigma for Ixx

@@

data at more orientation angles could improve the accuracy further and
is generally possible with standard diamond frame bicylces.

@@

    need sigma

@@

Calculating all three $J_{i}$ allows one to form

@@

and the unknown global inertia tensor can be solved for.

@@

        \includegraphics[width=2.75in]{../../../images/YellowFrameCompoundFirst.jpg}

@@

        \includegraphics[width=2.75in]{../../../images/BrowserInsForkCompoundFirst.jpg}

@@

        \caption{\subref{fig:frameCompound} The rear frame hung as a compound
        pendulum. \subref{fig:forkCompound} The Browser fork hung as a

@@

\ref{eq:inRel} and the compound is calculated with

@@

@@

    \caption{\subref{fig:StratosFork} The Stratos fork and handlebar assembly hung as a torsional
    \label{fig:torsional}

@@

Once all of the bicycle parameters have been calculated the canonical matrices
can be formed and the linear dynamics of the bicycles can be explored. We also

@@

rider was assumed to be in same position and posture for each bicycle relative

@@

\begin{table}[tbp]

@@

    \caption{Mass, center of mass and moment of inertia for the Person relative

@@

    \caption{The eigenvalues verus speed for all eight bicycles.}

@@

The eigenvalues can be plotted versus forward speed.

@@

characteristics of the Whipple bicycle model: four real roots at very slow
speeds, two of which are unstable; a complex pair that is unstable and lower
speeds and stable at higher speeds; and a root that is sligthly unstable at
higher speeds. The one noticable difference is that the capsize and caster
modes colasece into a complex pair between about 0.5 and 3 m/s. The frequency
of oscillation is similar to that of the weave mode. The root locus in the real

@@

roll with a 90 degree phase, both of their magitudes being similar,

@@

    \caption{The eigenvector components for the mode with the second complex
    pair at low speeds.}

@@

typical characteristics of the Whipple model. Reversing the fork on the yellow
bike lowers the weave critical speed. Also the addition of weight to the rear
rack of the Browser does blank.

@@

    \caption{The eigenvalues verus speed for all eight bicycles with the same
    rider.}

@@

The frequency response of the bicycles and bicycle with rider also reveal some
interesting things. In the steer torque to roll Bode diagram the magnitude
diffence among bicycles can vary up to 10 dB (0.148 radians per newton meter of
torque). The difference in the frequency response for the bicycle with the

@@

Fig.~\ref{fig:bikeRiderBode}, as the rider's mass and
inertia play a bigger roll.

@@

    \caption{The eigenvalues verus speed for all eight bicycles with the same
    rider.}

@@

    \caption{The eigenvalues verus speed for all eight bicycles with the same
    rider.}

@@

We present a detailed method of accurately estimating the physical parameters
of a bicycle need for the benchmarked Whipple bicycle
model~\cite{Meijaard2007}. Then we measure eight different bicycles providing
both the raw data, parameter sets and linear model coefficient matrices with
the uncertainties for each value. Then we present a brief comparison of the
eight bicycles using eigenanalysis and Bode analysis.

@@

This work was partially funded by the National Science Foundation.

\bibliographystyle{acm}
\bibliography{bicycle}
\appendix
%\section{RAW DATA}
%\begin{landscape}
%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
%\section{PARAMETERS}
%\subsection{BICYCLE}
\begin{landscape}
\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
%\subsection{BICYCLE WITH RIDER}
%\begin{landscape}
\subfile{../../../tables/BikeRider/Parameters/MasterParTable.tex}
%\end{landscape}
%\section{CANONICAL MATRICES}
%\subsection{BICYCLE}
%\begin{landscape}
\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
%\end{landscape}
%\subsection{BICYCLE WITH RIDER}
%\begin{landscape}
\subfile{../../../tables/BikeRider/Canonical/MasterCanTable.tex}
\end{landscape}
\end{document}

>>>>

\begin{center}
\begin{tabular}{c}
$^*$ Department of Mechanical and Aerospace Engineering\\
University of California, Davis\\
One Shields Avenue, Davis, CA, 95616, USA\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu\\
\end{tabular}
\begin{tabular}{c}
$^\dag$ Laboratory for Engineering Mechanics\\
Delft University of Technology\\
Mekelweg 2, 2628CD Delft, The Netherlands\\
e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com\\
\end{tabular} \\ \vspace{2.5ex}
\end{center}
\section*{ABSTRACT}
Accurate measurements of a bicycle's physical parameters are required for
realistic dynamic simulations and analysis. The most basic models require the
geometry, mass, mass location and mass distributions for the rigid bodies. More
complex models require estimates of tire characteristics, human
characteristics, friction, stiffness, damping, etc. In this
paper we present the measurement of the minimal bicycle parameters required for
the benchmark Whipple bicycle model presented in~\cite{Meijaard2007}. This
model is composed of four rigid bodies, has ideal rolling and frictionless joints,
and is laterally symmetric. A set of 25 parameters describes the
geometry, mass, mass location and mass distribution of each of the rigid
bodies. The experimental methods used to estimate the parameters described
herein are based primarily on the work done in~\cite{Kooijman2006} but have
been refined for improved accuracy and methodology. Koojiman's work was
preceded by \cite{Roland1971} who measured a bicycle in a
similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who used
similar techniques with scooters.

@@

These eight different parameter sets can be used with, but are not
limited to, the benchmark bicycle model. The accuracy of all the measurements
are presented up through the eigenvalue prediction of the linear model. The
accuracies are based on error propagation theory with correlations taken into
account.

@@

\section{INTRODUCTION}
This work is intended to document the indirect measurement of eight real
bicycles' physical parameters. The physical parameters measured are those
needed for the benchmark Whipple bicycle model presented in
\cite{Meijaard2007}. The work is based on techniques used to measure the
instrumented bicycles in \cite{Kooijman2006}, \cite{Kooijman2008} and
\cite{Moore2009a}. We improve upon these methods by both increasing and reporting
the accuracies of the measurements and by measuring the complete moments of
inertia of the frame and fork needed for analysis of the nonlinear model.
Furthermore, very little data exists on the physical parameters of different
types of bicycles and this work aims to provide a small sample of bicycles.

D\"{o}hring~\cite{Dohring1953} and Singh and Goel~\cite{Singh1971} measured the
physical parameters of scooters.
Roland and Massing~\cite{Roland1971} measured the physical parameters of a
bicycle in much the
same way as is presented, including calculations of uncertainty from the indirect
measurement techniques. Patterson~\cite{Patterson2004} used a swing to measure the
inertia of a bicycle and rider. The present work is based on the work
done by Kooijman~\cite{Kooijman2006} using much of the same apparatus and
refining the measurement technique.

@@

where the entries of the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
matrices are combinations of 25 bicycle physical parameters that

@@

necessarily a minimum set for the Whipple model, as shown in \cite{Sharp2008},
but are useful as they represent more intuitively measurable quantities.

@@

Fig.~\ref{fig:bicycles}. The three Batavus bicycles were donated by the

@@

variety, a mountain and road bike. The yellow bike is used to demonstrate

@@

    \item[Batavus Browser (B, B*)]{The Batavus Browser Fig.~\ref{fig:browser}
        is an average priced Dutch city

@@

        bars for an upright posture and includes various accessories for utility
        purposes. We measured the physical proprieties of the stock Browser model and

@@

        Dutch city bike for touring. It has an aluminum frame, an
        eight speed internal hub, upright handlebars, accessories for utility

@@

        lightweight front suspended mountain bike. It has a aluminum frame, large low
        pressure mountain bike tires, and is for racing with few extra accessories.}

@@

        aluminum. It has a seven-speed internal hub and mountain

@@

        accessories for utility such as a rear rack, fenders, light and

@@

    \item[Yellow Bicycle (Y, Y*)]{The yellow bicycle Fig.~\ref{fig:yellow} is used in the lab to

@@

        aluminum road frame of unknown make with the most of components

@@

        are the only parts on the bike. This bicycle was measured with both the
        fork in normal position and reversed. The fork was reversed to

@@

The 25 parameters can be estimated using many techniques. Where possible we

@@

We took great care to improve and report the accuracy of the
measurements of the parameters. Following the thrust of

@@

the measurements, $u,v,\ldots$, then $x$ is a random variable defined as

@@

Using the definitions for variance and covariance,

@@

hereafter have uncorrelated variables but a few do not and the covariance has to

@@

The radii of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were

@@

tape measure (resolution: 2mm) was pulled tight and taped on a flat level smooth floor. The tire

@@

$\pm0.2^{\circ}$ accuracy. The bicycle frame was fixed perpendicular to the
ground, the steering angle was set to the nominal, tire pressures were at

@@

fork blades were aligned such that the axle axis was parallel to the table

@@

We measured the wheelbase with the bicycle in nominal
configuration described in Sec.~\ref{sec:headtube}. We used a tape measure

@@

    \caption{\subref{fig:level} The digital level was mounted to a straight edge aligned

@@

orientations Fig.~\ref{fig:triangle}. The slope can be shown to be

@@

Theoretically, the center of mass lies on each line but due to experimental
error, if there are more than two lines, the lines do not cross all at the same point. Only two lines
are required to calculate the center of mass of the laterally

@@

Alternatively, the three lines can be treated as an over determined linear

@@

solution is not the same as the triangle centroid method.

@@

The solution with the higher accuracy is the preferred one.

@@

compound and torsional pendulums and measuring their periods of oscillation when
perturbed at small angles. The rate of oscillation was measured

@@

\textsc{Matlab}. The measurement durations were either 15 or 30 secs and each moment of

@@

determine only the period of oscillation which is needed for the moment of inertia

@@

The function Eqn~\ref{eqn:decayOs} was fit to the data using a nonlinear least squares fit
routine for each experiment to determine the quantities $A$, $B$, $C$, $\zeta$,
and $\omega$.

@@

The physical phenomenon observed corresponding to data sets such as these was that
the bicycle frame or fork was perturbed torsionally. After set into motion the
torsional motion died out and a longitudinal swinging motion increased. The
motions alternated back and forth with neither ever reaching zero. The
frequencies of these motions were very close to one another and it is not
apparent how dissect the two. We explored fitting to a function such as
\begin{equation}
    f(t) = A\sin{(\omega_1 t)} + B\sin{(\omega_2 t + \phi)} + C
    \label{eqn:sumSines}
\end{equation}
But the fit predicts that $\omega_1$ and $\omega_2$ are very similar
frequencies. There was no easy way to choose which of the two $\omega$'s was
the one associated with the torsional oscillation. Some work was done to model
the torsional pendulum as a laterally flexible beam to determine this, but we
thought accuracy of the period calculation would not
improve enough for the effort required. Future experiments should simply
prevent the swinging motion of the pendulum without damping the torsional
motion.

The period for a damped oscillation is
\begin{equation}
    T = \frac{2\pi}{\sqrt{1-\zeta^2}\omega_n}
    \label{eqn:periodDamped}
\end{equation}
The uncertainty in the period, $T$, can be determined from the fit. Firstly,
the variance of the fit is
\begin{equation}
    \sigma_y^2 =
    \frac{1}{N-5}\sum_{i=1}^N(y_{mi}-\bar{y}_m)^2-(y_{pi}-\bar{y}_m)^2
    \label{eqn:fitVariance}
\end{equation}
The covariance matrix of the fit function can be formed
\begin{equation}
    \mathbf{U} = \sigma_y^2\mathbf{H}^{-1}
    \label{eqn:covariance}
\end{equation}
where $\mathbf{H}$ is the Hessian~\cite{Hubbard1989b}. $\mathbf{U}$ is a $5\times5$ matrix with the
variances of each of the five fit parameters along the diagonal.
The variance of $T$ can be computed using the variance of $\zeta$ and $\omega$. It
is important to note that the uncertainties in the period are very low
($<1e-4$), even for the fits with low $r^2$ values.
\subsection{TORSIONAL PENDULUM}
A torsional pendulum was used to measure all moments of inertia about axes
in the laterally symmetric plane of each of the wheels, fork and frame. The
pendulum is made up of a rigid mount, an upper clamp, a torsion rod, and
various lower clamps.
\begin{figure}[tb]
    \begin{center}
        \includegraphics[width=2in]{../../../images/fixture.jpg}
    \end{center}
    \caption{The rigid pendulum fixture mounted to a concrete column.}
    \label{fig:fixture}
\end{figure}
A 5 mm diameter, 1 m long mild steel rod was used as the torsion
spring. A lightweight, low relative moment of inertia clamp was constructed
that could clamp the rim and the tire. The moments of inertia of the clamps
were neglected. The wheel was hung freely such that the
center of mass aligned with the torsional pendulum axis and then
secured. The wheel was then perturbed and oscillated about the pendulum
axis. The rate gyro was mounted on the clamp oriented along the pendulum
axis.

@@

inertia Fig.~\ref{fig:rod}. A torsional pendulum almost identical to the one used in
\cite{Kooijman2006} was used to measure the average period $\overline{T}_i$ of

@@

are assumed symmetric about three orthogonal planes so products of
inertia are zero. The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring

@@

torsional pendulum setup and Eq.~\ref{eq:torPend}.The wheels are assumed to be
laterally symmetric and about any radial axis. Thus only two
moments of inertia are required for the set of benchmark parameters. The moment

@@

gyro was attached to the spokes near the hub and oriented mostly along the

@@

axis which added a very low frequency component of rate along the vertical
radial axis, but this should not affect the period estimation about the compound
pendulum axis. A fixture that prevented precession would be preferable for

@@

inertia about the axle is calculated from:

@@

%\begin{equation}
    %need the sigma for IRyy
%\end{equation}

@@

the torsional pendulum axis and then secured. The wheel was then
perturbed and oscillated about the vertical pendulum axis. The radial moment of inertia can can calculated as such:

@@

%\begin{equation}
    %need the sigma for Ixx
%\end{equation}

@@

data at more orientation angles could improve the accuracy and
is generally possible with standard diamond frame bicycles.

@@

%\begin{equation}
    %need sigma
%\end{equation}

@@

Similarly, calculating all three $J_{i}$ allows one to form

@@

and the moments of inertia can be solved for.

@@

        \includegraphics[width=2in]{../../../images/YellowFrameCompoundFirst.jpg}

@@

        \includegraphics[width=2in]{../../../images/BrowserInsForkCompoundFirst.jpg}

@@

        \caption{\subref{fig:frameCompound} Rear frame hung as a compound
        pendulum. \subref{fig:forkCompound} Browser fork hung as a

@@

\ref{eq:inRel} and the compound pendulum calculations is calculated with

@@

    \caption{The Stratos fork and handlebar assembly hung as a torsional

@@

@@

Once all bicycle parameters have been calculated the canonical matrices
can be formed and the linear dynamics of the bicycles can be explored. The
values of the canonical matrices can be found in the second table in Appendix~\ref{sec:partables}. We also

@@

rider was assumed to be in the same position and posture for each bicycle relative

@@

\begin{table}[tb]

@@

    \caption{Mass, center of mass and moment of inertia for the rider relative

@@

    \caption{Eigenvalues versus speed for all eight bicycles without the rider.}

@@

The eigenvalues of the bicycles with (Fig.~\ref{fig:bikeEigPlot}) and without
(Fig.~\ref{fig:bikeRiderEigPlot} the rider can be plotted versus forward speed.

@@

characteristics of the benchmark bicycle: four real roots at very slow
speeds, two of which are unstable; a complex pair that is unstable at lower
speeds and stable at intermediate speeds; and a root that is mildly unstable at
higher speeds. The one noticeable difference is that the capsize and caster
modes are contained in a complex pair between about 0.5 and 3 m/s. The frequency
of oscillation is of comparable magnitude to that of the weave mode. But, the root locus in the real

@@

roll with a 90 degree phase, both of their magnitudes being similar,

@@

    \caption{Eigenvector components for the second complex mode
    pair at low speed.}

@@

typical characteristics of the benchmark bicycle model. Reversing the fork on the yellow
bike lowers the weave critical speed and increases the stable speed range.
Also, the addition of weight to the rear rack of the Browser does little to the
eigenvalues.

@@

    \caption{Eigenvalues versus speed for all eight bicycles with the same
    rigid rider.}

@@

The frequency response of the bicycles (Fig.~\ref{fig:bikeBode}) and bicycle
with rider (Fig.~\ref{fig:bikeRiderBode}) also reveal some
interesting things. In the steer-torque-to-roll Bode diagram the magnitude
difference among bicycles can vary up to 10 dB (or about 8.5 degrees per Newton-meter of
torque) for the particular speed shown. The difference in the frequency response for the bicycle with the

@@

Fig.~\ref{fig:bikeRiderBode}, as the rider's mass and inertia play a larger roll.

@@

    \caption{The frequency response for steer-torque-to-roll for all eight
    bicycles without the rider at 2 m/s.}

@@

    \caption{The frequency response for steer-torque-to-roll for all eight
    bicycles with the same rigid rider at 2 m/s.}

@@

We have presented a detailed method to accurately estimate the physical parameters
of a bicycle needed for the benchmarked Whipple bicycle
model~\cite{Meijaard2007}. We measured eight different bicycles providing
both the parameter sets and linear model coefficient matrices
for the bicycles alone and the bicycles with the same
rigid rider. The uncertainties in the parameters and matrix coefficients are
included for the bicycle alone. Finally, we have presented a brief comparison of the
eight bicycles using eigenanalysis and Bode frequency response.

@@

This material is based upon work partially supported by the National Science
Foundation under Grant No. 0928339.

\bibliographystyle{acm}
\bibliography{bicycle}
\appendix
\section{PARAMETER TABLES}
\label{sec:partables}
The tabulated values for the both the physical parameters and the canonical matrix
coefficients are shown in the following four tables. The uncertainties in the
estimations of both the parameters and coefficients are also shown for the
bicycle without a rider.
%\section{RAW DATA}
%\begin{landscape}
%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
%\section{PARAMETERS}
%\subsection{BICYCLE}
\begin{landscape}

\begin{table}[tb]
\caption{The parameters for the eight bicycles with uncertainties in the
estimations.}
\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
\label{tab:bicyclePar}
\end{table}

\begin{table}[tb]
\caption{The canonical matrix coefficients for the eight bicycles with the
uncertainty in the estimations.}
\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
\label{tab:bicycleCan}
\end{table}

\begin{table}[tb]
\caption{The parameters for the eight bicycles with the same rigid rider.}
\subfile{../../../tables/BikeRider/Parameters/MasterParTable.tex}
\label{tab:bicycleRiderPar}
\end{table}

\begin{table}[tb]
\caption{The canonical matrix coefficients for the eight bicycles with the
rigid rider.}
\subfile{../../../tables/BikeRider/Canonical/MasterCanTable.tex}
\label{tab:bicycleRiderCan}
\end{table}

\end{landscape}
\end{document}
